### 2012-09-04 sections/evaluation.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * performance optimizations

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{filter}
    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
    \label{fig:benchmarks-filter}
  \end{center}
\end{figure}

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then again SilK and \texttt{nfdump} also use their own file
format. As a result, the amount of data that needs to read (or written) may be
different to what it is for \ac{NFQL} and \texttt{flowtools}.  On the other
hand, \texttt{nfdump} appears to be significantly faster than the rest. This
is because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
that are read and written used \texttt{lzo} compression scheme which trades
space for achieving faster compression and decompression. It is important to
note, that all the tools were single-threaded in this evaluation, and did not
completely utilize the $24$ available cores. It comes as a realization, that
adding \texttt{lzo} compression will drastically improve \ac{NFQL}'s filter
performance.

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{grouper}
    \caption{Grouper Stage: NFQL vs SiLK}
    \label{fig:benchmarks-grouper}
  \end{center}
\end{figure}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{groupfilter}
    \caption{Group Filter Stage: NFQL vs SiLK}
    \label{fig:benchmarks-groupfilter}
  \end{center}
\end{figure}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{merger}
    \caption{Merger Stage}
    \label{fig:benchmarks-merger}
  \end{center}
\end{figure}

@@

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{ungrouper}
    \caption{Ungrouper Stage}
    \label{fig:benchmarks-ungrouper}
  \end{center}
\end{figure}

>>>>

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{filter}
    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
    \label{fig:benchmarks-filter}
  \end{center}
\end{figure}

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then again SilK and \texttt{nfdump} also use their own file
format. As a result, the amount of data that needs to read (or written) may be
different to what it is for \ac{NFQL} and \texttt{flowtools}.  On the other
hand, \texttt{nfdump} appears to be significantly faster than the rest. This
is because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
that are read and written used \texttt{lzo} compression scheme. It is
important to note, that all the tools were single-threaded in this evaluation,
and did not completely utilize the $24$ available cores. It comes as a
realization, that adding \texttt{lzo} compression will drastically improve
\ac{NFQL}'s filter performance.

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{grouper}
    \caption{Grouper Stage: NFQL vs SiLK}
    \label{fig:benchmarks-grouper}
  \end{center}
\end{figure}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{groupfilter}
    \caption{Group Filter Stage: NFQL vs SiLK}
    \label{fig:benchmarks-groupfilter}
  \end{center}
\end{figure}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{merger}
    \caption{Merger Stage}
    \label{fig:benchmarks-merger}
  \end{center}
\end{figure}

@@

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{ungrouper}
    \caption{Ungrouper Stage}
    \label{fig:benchmarks-ungrouper}
  \end{center}
\end{figure}

### 2012-09-04 sections/implementation/adaptable-compression-levels.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)
* revision with professor's comments   * performance optimizations

The flow-records echoed to the standard output can also be written to a
Netflow $v5$ \texttt{flow-tools} file. The \texttt{--dirpath} switch allows
one to provide a directory path where the results can be stored.  Each stream
is stored as its own file with an ID to disambiguate it. Results from each
stage of the pipeline can also be written to separate files with the increase
in the verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose}
work well together to adjust the level at which the writes are to be made.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{zlevel}
    \caption{F$(v2)$: \texttt{z-level} Effect on Performance}
    \label{fig:engine-zlevel}
  \end{center}
\end{figure}

The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
the results written to the \texttt{flow-tools} files. \texttt{zlib} supports
$9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
supports \texttt{--zlevel} switch to allow the user to supply its desired
choice of the of the compression level. A default level of $5$ is used for
writes if the switch is not supplied during runtime.  Fig.
\ref{fig:engine-zlevel} shows the time taken to write a sample of records
passing the filter stage for each \texttt{z-level}. It goes to show that each
level adds its own performance overhead and must be used with discretion.

>>>>

The flow-records echoed to the standard output can also be written to Netflow
$v5$ \texttt{flow-tools} files. The \texttt{-{}-dirpath} switch allows one to
provide a directory path where the results can be stored. Results from each
stage of the pipeline can also be written to separate files with the increase
in the verbosity level. In fact, \texttt{-{}-dirpath} and \texttt{-{}-verbose}
work well together to adjust the level at which the writes are to be made.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{zlevel}
    \caption{\texttt{z-level} Effect on Performance}
    \label{fig:engine-zlevel}
  \end{center}
\end{figure}

The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
the results written to the \texttt{flow-tools} files. \texttt{zlib} supports
$9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
supports \texttt{-{}-zlevel} switch to allow the user to supply its desired
choice of the of the compression level. A default level of $5$ is used for
writes if the switch is not supplied during runtime.  Fig.
\ref{fig:engine-zlevel} shows the time taken to write a sample of records
passing the filter stage for each \texttt{z-level}. It goes to show that each
level adds its own performance overhead and must be used with discretion. It
is also important to note that other tools may use different compression
algorithms. \texttt{nfdump} for instance, uses \texttt{lzo} \cite{lzo}
compression to trade space for faster compression and decompression.

### 2012-09-04 sections/implementation/performance-optimizations.tex
* revision with professor's comments   * performance optimizations

\ac{NFQL} has dedicated comparator functions for each type of operation and
the type of the field offset it operates upon. It is not guaranteed that given
the type of the query and trace, the program will eventually go through each
stage of the pipeline. It is also possible that the program exits before,
because there is nothing more for the next stage to compute. The function
pointers are therefore set as late as possible and are are called from their
respective stages just before the comparison is needed.  As a result, we save
the computation time wasted in setting the function pointer for stage if it is
is never executed.

\ac{NFQL} pushes the filter stage out of the branch and closer to where the
trace is originally read. This reduces the memory allocation to only records
that passed the filter stage thereby reducing the memory footprint of the
execution engine.

@@

The results from each stage of the pipeline are echoed to the standard output
just before the execution engine exits. This leads to an additional loop to
echo the results, however echoes to the standard output are only used for
debugging purposes. The writes to a file can however be legitimately requested
in a real network analysis task. As a result, it is essential to avoid
additional loops when performing writes to a file. The execution engine,
therefore writes each result record to a file as soon as it is seen by the
pipeline stage.

>>>>

\ac{NFQL} has dedicated comparator functions for each type of operation and
the type of the field offset it operates upon. It is not guaranteed that given
the type of the query and trace, the program will eventually go through each
stage of the pipeline. It is also possible that the program exits before,
because there is nothing more for the next stage to compute. The function
pointers are therefore set as late as possible and are called from their
respective stages just before the comparison is needed.  As a result, we save
the computation time wasted in setting the function pointer for stage if it is
is never executed.

\ac{NFQL} pushes the filter stage out of the branches and closer to where the
trace is originally read. This reduces the memory allocation to only records
that passed the filter stage, thereby reducing the memory footprint of the
execution engine.

@@

The results from each stage of the pipeline are echoed to the standard output
just before the execution engine exits. This leads to an additional loop to
echo the results, however echoes to the standard output are only used for
debugging purposes. The writes to a file can, however, be legitimately
requested in a real network analysis task. As a result, it is essential to
avoid additional loops when performing writes to a file. The execution engine
therefore writes each result record to a file as soon as it is seen by the
pipeline stage.
